---
title: "Ontario COVID-19 Reporting Dashboard"
author: Patrick Reza Schnurbusch 
date: '`r paste("Updated: ", format(Sys.time(), "%Y-%m-%d %H:%M %Z"))`'
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---

<b> A dashboard to display and track COVID-19 pandemic data in Ontario </b>

```{r setup, include=FALSE}
library(flexdashboard)
library(readr)
library(dplyr)
library(plotly)
library(DT)
library(tidyverse)

library(ggplot2)
library(plyr)
library(scales)
library(zoo)
library(tidyquant) 

# urlfile="https://data.ontario.ca/dataset/f4f86e54-872d-43f8-8a86-3892fd3cb5e6/resource/ed270bb8-340b-41f9-a7c6-e8ef587e6d11/download/covidtesting.csv"
# 
# caseStatusOntario <-read.csv(url(urlfile))
# 
# #data dict URL 
# #https://data.ontario.ca/dataset/status-of-covid-19-cases-in-ontario
# 


#pull .csv() for today by using the first 10 character for filtering

dirPath <- "../../MyDirectory../" #put your current directory here 

target_files <- as.data.frame(list.files(dirPath))
names(target_files) <- "files"
target_files$download_date <- as.Date(substr(target_files$files, 1, 10))

target_files <- target_files %>% 
  dplyr::arrange(download_date) %>% 
  dplyr::slice_tail()


#now we have 1 row that has the most recent date and the location of the target 
dirPath <- paste0("../../Datasets/..", target_files$files)

caseStatusOntario <-read.csv(file = paste(dirPath))

caseStatusOntario$Reported.Date <- as.Date(caseStatusOntario$Reported.Date)
```

TESTING
=====================================  


Row {data-width=1000}
-----------------------------------------------------------------------

### <b> Daily Confirmed Positives </b>

```{r}
fig <- plot_ly(x = caseStatusOntario$Reported.Date, 
               y = caseStatusOntario$Confirmed.Positive, 
               type = 'scatter', mode = 'lines', name = 'Confirmed Positive', 
               fill = 'tozeroy',
        fillcolor = '#2858a6',
        line = list(width = 0.5))

fig <- fig %>% layout(title = "",
                      yaxis = list(title = 'Tested Positive'),
                      xaxis = list(title = 'Reported Date'),
                      hovermode = "x unified")

fig
```

Row {data-width=1000}
-----------------------------------------------------------------------
### <b> Positives vs Total Tested </b>
```{r}
fig <- plot_ly(caseStatusOntario, x = ~Reported.Date, y = ~Confirmed.Positive, type = 'bar', name = 'Confirmed Positives', marker = list(color = 'FF5959'))
fig <- fig %>% add_trace(y = ~Total.tests.completed.in.the.last.day, name = 'Total Tested', marker = list(color = '6EC5E9'))
fig <- fig %>% layout(title = "Positives vs Total Tested", 
                      yaxis = list(title = 'Count'), barmode = 'stack', 
                      hovermode = "x unified")

fig

```


Column {data-width=500}
-----------------------------------------------------------------------

### <b> Test Positive Rate </b> 

```{r}
test.positive.rate <- caseStatusOntario %>% 
  dplyr::select(Reported.Date, Confirmed.Positive, Total.tests.completed.in.the.last.day) %>% 
  dplyr::mutate(Test.positive.rate = round(Confirmed.Positive/Total.tests.completed.in.the.last.day*100,2))

fig <- plot_ly(test.positive.rate, x = ~Reported.Date, y = ~Test.positive.rate, type = 'bar', name = 'Test Positive Rate', 
               marker = list(color = '#10fbc9'))
fig <- fig %>% layout(title = "", 
                      yaxis = list(title = 'Test Positive (%)'),
                      hovermode = "x unified")

fig
```















CASE STATUSES
===================================== 
Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### <b> Linear Scale </b>
```{r}
fig <- plot_ly(x = ~caseStatusOntario$Reported.Date, y = ~caseStatusOntario$Total.Cases, type = 'scatter', mode = 'none', name = 'Total Cases', fill = 'tozeroy',
               fillcolor = '#FFC56C')
fig <- fig %>% add_trace(x = ~caseStatusOntario$Reported.Date, y = ~caseStatusOntario$Resolved, name = 'Resolved', fill = 'tozeroy',
                         fillcolor = '#6EC5E9')
fig <- fig %>% add_trace(x = ~caseStatusOntario$Reported.Date, y = ~caseStatusOntario$Deaths, name = 'Deaths', fill = 'tozeroy',
                         fillcolor = '#FF5959')

fig <- fig %>% layout(title = "Daily Cases by Status", 
                      xaxis = list(title = ''),
                      yaxis = list(title = 'Cumulative Case Counts'), 
                      hovermode = "x unified")

fig <- fig %>% 
  layout(xaxis = list(rangeselector = list(
    buttons = list(
      
      list(
        count = 1,
        label = "1 mo",
        step = "month",
        stepmode = "backward"),
      
      list(
        count = 3,
        label = "3 mo",
        step = "month",
        stepmode = "backward"),
      
      list(
        count = 6,
        label = "6 mo",
        step = "month",
        stepmode = "backward"),
      
      list(
        count = 1,
        label = "1 yr",
        step = "year",
        stepmode = "backward"),
      
      list(
        count = 1,
        label = "YTD",
        step = "year",
        stepmode = "todate"),
      list(step = "all")))
    #,rangeslider = list(type = "date")
    ))

fig
```

### <b> Logarithmic Scale </b>
```{r}
fig <- fig %>% layout(yaxis = list(type = "log"))

fig <- fig %>% layout(title = "Daily Cases by Status",
                      xaxis = list(title = "Daily Cases by Status (Log)"),
                      yaxis = list (title = "Cumulative Case Counts (Log)"))
fig

```


Row {.tabset .tabset-fade}
-------------------------------------

### <b> New Cases (Relative Change) </b>

```{r}
#plotting the percentage increase for each case 
case.percentages <- caseStatusOntario %>% 
  dplyr::select(Reported.Date, Total.Cases, Resolved, Deaths)
 
case.percentages <- case.percentages %>%
  mutate(New_Total.Cases = Total.Cases - lag(Total.Cases,1, default = 0.0), 
         Total.Cases.Change.Pct = round(New_Total.Cases/Total.Cases*100 ,2), 
         
         New_Resolved = Resolved - lag(Resolved,1, default = 0.0), 
         Resolved.Change.Pct = round(New_Resolved/Resolved*100 ,2), 
         
         New_Deaths = Deaths - lag(Deaths,1, default = 0.0), 
         Deaths.Change.Pct = round(New_Deaths/Deaths*100 ,2))
```


```{r}
fig <- plot_ly(case.percentages, x = ~Reported.Date, y = ~Total.Cases.Change.Pct, type = 'bar', 
               name = 'New Total Cases', 
               marker = list(color = '#FFC56C'))

fig <- fig %>% add_trace(y = ~Resolved.Change.Pct, 
                         name = 'New Resolved', 
                         marker = list(color = '#6EC5E9'))

fig <- fig %>% add_trace(y = ~Deaths.Change.Pct, 
                         name = 'New Deaths', 
                         marker = list(color = '#FF5959'))

fig <- fig %>% layout(title = "", 
                      xaxis = list(title = ''),
                      yaxis = list(title = 'Relative Change (%)'), 
                      barmode = 'stack', 
                      hovermode = "x unified")
fig
```


### <b> New Cases (Absolute Change) </b>
    
```{r}
fig <- plot_ly(case.percentages, x = ~case.percentages$Reported.Date, y = ~case.percentages$New_Total.Cases, type = 'bar', 
               name = 'New Total Cases', 
               marker = list(color = '#FFC56C'))

fig <- fig %>% add_trace(y = ~case.percentages$New_Resolved, 
                         name = 'New Resolved',
                         marker = list(color = '#6EC5E9'))

fig <- fig %>% add_trace(x = ~case.percentages$Reported.Date, y = ~case.percentages$New_Deaths, 
                         name = 'New Deaths', 
                         marker = list(color = '#FF5959'))

fig <- fig %>% layout(title = "", 
                      xaxis = list(title = ''),
                      yaxis = list(title = 'Absolute Change (n)'), 
                      barmode = 'stack',
                      hovermode = "x unified")

fig
```

Row 
-------------------------------------
    
### <b> New Cases vs. New Recoveries </b>
    
```{r}
newCases_vs_newRecoveries <- caseStatusOntario %>% 
  dplyr::select(Reported.Date, Confirmed.Positive, Resolved) %>% 
  dplyr::mutate(NewCases = Confirmed.Positive - lag(Confirmed.Positive,1, default = 0.0), 
                NewRecoveries = Resolved - lag(Resolved,1, default = 0.0))
  
fig <- plot_ly(data = newCases_vs_newRecoveries, 
               x = ~Reported.Date, y = ~NewCases, type = 'scatter', 
               mode = 'lines', name = 'New Cases', 
               line = list(color = 'FF5959', width = 2.5))

fig <- fig %>% add_trace(x = ~newCases_vs_newRecoveries$Reported.Date, y = ~newCases_vs_newRecoveries$NewRecoveries, 
                         name = 'New Recoveries', mode = 'lines',
                         line = list(color = '1CE8A9', width = 2.5))

fig <- fig %>% layout(title = "", 
                      xaxis = list(title = ''),
                      yaxis = list(title = 'New Daily COVID-19 Cases + Cured Cases'), 
                      hovermode = "x unified")
fig
```



Row {.tabset .tabset-fade}
-------------------------------------

    
```{r}
#hospitalizations 
hospitalizations <- caseStatusOntario %>% 
  dplyr::select(Reported.Date, 
                Number.of.patients.hospitalized.with.COVID.19, 
                Number.of.patients.in.ICU.with.COVID.19, 
                Number.of.patients.in.ICU.on.a.ventilator.with.COVID.19) %>% 
  
  dplyr::mutate(
    #absolute changes 
                New_hospitalized = Number.of.patients.hospitalized.with.COVID.19 - lag(Number.of.patients.hospitalized.with.COVID.19,1, default = 0.0),
                New_ICU = Number.of.patients.in.ICU.with.COVID.19 - lag(Number.of.patients.in.ICU.with.COVID.19,1, default = 0.0), 
                New_ventilator = Number.of.patients.in.ICU.on.a.ventilator.with.COVID.19 - lag(Number.of.patients.in.ICU.on.a.ventilator.with.COVID.19,1, default = 0.0),
                
                #relative changes 
                
                New_hospitalized.pct = round(New_hospitalized/Number.of.patients.hospitalized.with.COVID.19*100 ,2), 
                New_ICU.pct = round(New_ICU/Number.of.patients.in.ICU.with.COVID.19*100 ,2), 
                New_ventilator.pct = round(New_ventilator/Number.of.patients.in.ICU.on.a.ventilator.with.COVID.19*100 ,2))

```

### <b> Current Hospitalizations </b>
```{r}
fig <- plot_ly(hospitalizations, x = ~Reported.Date, y = ~Number.of.patients.hospitalized.with.COVID.19, 
               type = 'scatter', 
               mode = 'markers',
               name = 'Patients Hospitalized', 
               marker = list(color = '#2C82C9', size = 10))

fig <- fig %>% add_trace(y = ~Number.of.patients.in.ICU.with.COVID.19, name = 'Intensive Care Patients', 
                         marker = list(color = '#FCB941', size = 10))

fig <- fig %>% add_trace(y = ~Number.of.patients.in.ICU.on.a.ventilator.with.COVID.19, name = 'ICU Ventilator Patients', 
                         marker = list(color = '#FC6042', size = 10))

fig <- fig %>% layout(title = "",
                      xaxis = list(title = ''), 
                      yaxis = list(title = 'Number of Patients'), barmode = 'stack', 
                      hovermode = "x unified")
fig
```

### <b> Change in Current Hospitalizations (Absolute) </b>

```{r}
fig <- plot_ly(data = hospitalizations, x = ~Reported.Date,y = ~New_hospitalized, 
               type = 'scatter',
               mode = 'lines+markers', 
               name = "Patients Hospitalized", 
               marker = list(color = '#2C82C9'),
               line = list(color = '#2C82C9', width = 1)
) 

fig <- fig %>% add_trace(y = ~New_ICU, name = "Intensive Care Patients", mode = 'lines+markers',
                         marker = list(color = '#FCB941'),
                         line = list(color = '#FCB941', width = 1)
)

fig <- fig %>% add_trace(y = ~New_ventilator, name = "ICU Ventilator Patients", mode = 'lines+markers',
                         marker = list(color = '#FC6042'),
                         line = list(color = '#FC6042', width = 1)
)


fig <- fig %>% layout(title = "",
                      xaxis = list(title = ''), 
                      yaxis = list(title = 'Change in Patients (n)'), barmode = 'stack', 
                      hovermode = "x unified")
fig
```

### <b> Change in Current Hospitalizations (Relative) </b> 

```{r}
fig <- plot_ly(data = hospitalizations, x = ~Reported.Date,y = ~New_hospitalized.pct, 
               type = 'scatter',
               mode = 'lines+markers', 
               name = "Patients Hospitalized", 
               marker = list(color = '#2C82C9'), 
               line = list(color = '#2C82C9', width = 1)
               ) 

fig <- fig %>% add_trace(y = ~New_ICU.pct, name = "Intensive Care Patients", mode = 'lines+markers',
                         marker = list(color = '#FCB941'),
                         line = list(color = '#FCB941', width = 1)
                         )

fig <- fig %>% add_trace(y = ~New_ventilator.pct, name = "ICU Ventilator Patients", mode = 'lines+markers',
                         marker = list(color = '#FC6042'),
                         line = list(color = '#FC6042', width = 1)
                         )

fig <- fig %>% layout(title = "",
                      xaxis = list(title = ''),  
                      yaxis = list(title = 'Change in Patients (%)'), barmode = 'stack', 
                      hovermode = "x unified")
fig
```

RATES
===================================== 

### Pandemic Rates

```{r}

disease.rate <- caseStatusOntario %>% 
  dplyr::select(Reported.Date, Total.Cases, Deaths) %>% 
  dplyr::mutate(death.rates.per.hundtho = round(Deaths / 14750000 * 100000,3),
                death.rate.pct = round(death.rates.per.hundtho/1000,3), 
                New_Total.Cases = Total.Cases - lag(Total.Cases,1, default = 0.0),
                incidence.rate = round(New_Total.Cases / 14750000 * 100000,3), 
                prevalence.rate = round(Total.Cases / 14750000 * 100000,3))  
```

Row {.tabset .tabset-fade}
-------------------------------------
### <b> Death Rates per 100K </b> 
```{r}
fig <- plot_ly(data = disease.rate, 
               x = ~Reported.Date, 
               y = ~death.rates.per.hundtho,
               type = 'scatter', 
               mode = 'markers', 
               
               # Hover text:
               text = ~paste("Reported Date: ", Reported.Date, 
                             '<br> <b>Death Rate Capita: ', death.rates.per.hundtho, '<b>'),

               marker = list(size = 10, 
                             opacity = 0.9, 
                              color = '#FF5959'))

fig <- fig %>% layout(title = '',
                      yaxis = list(title = "Death Rate", zeroline = FALSE),
                      xaxis = list(title = " ", zeroline = FALSE),
                      hovermode = "x unified")

fig
```

### <b> Incidence Rates per 100K </b> 
```{r}
fig <- plot_ly(data = disease.rate, 
               x = ~Reported.Date, 
               y = ~incidence.rate,
               type = 'scatter', 
               mode = 'markers', 
               
               # Hover text:
               text = ~paste("Reported Date: ", Reported.Date, 
                             '<br> <b>Incidence Rate per Capita: ', incidence.rate, '<b>'),
               
               marker = list(size = 10, 
                             opacity = 0.9,
                             color = '#1CE8A9'))

fig <- fig %>% layout(title = '',
                      yaxis = list(title = "Incidence Rate", zeroline = FALSE),
                      xaxis = list(title = " ", zeroline = FALSE),
                      hovermode = "x unified")

fig
```

### <b> Prevalence Rates per 100K </b>
```{r}
fig <- plot_ly(data = disease.rate, 
               x = ~Reported.Date, 
               y = ~prevalence.rate,
               type = 'scatter', 
               mode = 'markers', 
               
               # Hover text:
               text = ~paste("Reported Date: ", Reported.Date, 
                             '<br> <b>Prevalence Rate per Capita: ', prevalence.rate, '<b>'),
               
               marker = list(size = disease.rate$prevalence.rate, 
                             opacity = 0.9, color = '#173F5F'))

fig <- fig %>% layout(title = '',
                      yaxis = list(title = "Prevalence.rate Rate", zeroline = FALSE),
                      xaxis = list(title = " ", zeroline = FALSE),
                      hovermode = "x unified")

fig
```

Row 
-------------------------------------

```{r}
disease.rate <- disease.rate %>% 
  dplyr::arrange(desc(Reported.Date))

datatable(disease.rate, 
          colnames = c('Reported Date', 'Total Cases', 'Deaths', 'Death Rate', 'Death Rate Percent', 'New Cases', 'Incidence Rate', 'Prevalence Rate'),
          class = 'cell-border stripe')
```


LONG TERM CARE HOMES
===================================== 
```{r}
ltc_cases <- caseStatusOntario %>% 
  dplyr::select("Reported.Date","Total.Positive.LTC.Resident.Cases",
                "Total.Positive.LTC.HCW.Cases",                            
                "Total.LTC.Resident.Deaths",
                "Total.LTC.HCW.Deaths") %>% 
  drop_na() %>% 
  
  #calc the absolute change (new numbers) 
  dplyr::mutate(New_LTC.Resident.Cases = Total.Positive.LTC.Resident.Cases - lag(Total.Positive.LTC.Resident.Cases,1, default = 0.0), 
                New_LTC.HCW.Cases = Total.Positive.LTC.HCW.Cases - lag(Total.Positive.LTC.HCW.Cases, 1, default = 0.0),
                New_LTC.Resident.Deaths = Total.LTC.Resident.Deaths - lag(Total.LTC.Resident.Deaths, 1, default = 0.0),
                New_LTC.HCW.Deaths = Total.LTC.HCW.Deaths - lag(Total.LTC.HCW.Deaths, 1, default = 0.0),
                #calculating the relative change as percentage
                
                New_LTC.Resident.Pct = round(New_LTC.Resident.Cases/Total.Positive.LTC.Resident.Cases * 100, 2),
                New_LTC.HCW.Pct = round(New_LTC.HCW.Cases/Total.Positive.LTC.HCW.Cases * 100, 2), 
                New_LTC.Resident.Deaths.Pct = round(New_LTC.Resident.Deaths/Total.LTC.Resident.Deaths * 100, 2),
                New_LTC.HCW.Deaths.Pct = round(New_LTC.HCW.Deaths/Total.LTC.HCW.Deaths * 100, 2)
    ) 
```

Row {.tabset .tabset-fade}
-------------------------------------
### <b> Daily Cases </b> 
```{r}
fig <- plot_ly(x = ~ltc_cases$Reported.Date, y = ~ltc_cases$Total.Positive.LTC.Resident.Cases, type = 'bar', 
               name = 'Residents', 
               marker = list(color = '#C7CEEA')) 

fig <- fig %>% add_trace(y = ~ltc_cases$Total.Positive.LTC.HCW.Cases, name = 'Health Care Workers', marker = list(color = '#FFB7B2')) 


fig <- fig %>% layout(title = "",
                      xaxis = list(title = "Time"),
                      yaxis = list (title = "Cumulative Case Counts"), 
                      barmode = 'stack',
                      hovermode = "x unified")
fig
```

### <b> New Cases (Relative Change) </b>

```{r}
fig <- plot_ly(x = ~ltc_cases$Reported.Date, y = ~ltc_cases$New_LTC.Resident.Pct, type = 'bar', 
               name = 'Residents', 
               marker = list(color = '#C7CEEA')) 

fig <- fig %>% add_trace(y = ~ltc_cases$New_LTC.HCW.Pct, name = 'Health Care Workers', marker = list(color = '#FFB7B2')) 


fig <- fig %>% layout(title = "",
                      xaxis = list(title = "Time"),
                      yaxis = list (title = "New Cases (%)"), 
                      barmode = 'stack',
                      hovermode = "x unified")
fig

```

### <b> New Cases (Absolute Change) </b>

```{r}
fig <- plot_ly(x = ~ltc_cases$Reported.Date, y = ~ltc_cases$New_LTC.Resident.Cases, type = 'bar', 
               name = 'Residents', 
               marker = list(color = '#C7CEEA')) 

fig <- fig %>% add_trace(y = ~ltc_cases$New_LTC.HCW.Cases, name = 'Health Care Workers', marker = list(color = '#FFB7B2')) 


fig <- fig %>% layout(title = "",
                      xaxis = list(title = "Time"),
                      yaxis = list (title = "New Cases"), 
                      barmode = 'stack',
                      hovermode = "x unified")
fig
```

Row {.tabset .tabset-fade}
-------------------------------------
### <b> Deaths </b>
```{r}
fig <- plot_ly(x = ~ltc_cases$Reported.Date, y = ~ltc_cases$Total.LTC.Resident.Deaths, type = 'bar', 
               name = 'Residents', 
               marker = list(color = '#AADAFF')) 

fig <- fig %>% add_trace(y = ~ltc_cases$Total.LTC.HCW.Deaths, name = 'Health Care Workers', marker = list(color = '#FF5251')) 


fig <- fig %>% layout(title = "",
                      xaxis = list(title = "Time"),
                      yaxis = list (title = "Cumulative Case Counts"), 
                      barmode = 'stack',
                      hovermode = "x unified")
fig
```

### <b> New Deaths per Day (Absolute Change) </b>
```{r}
fig <- plot_ly(x = ~ltc_cases$Reported.Date, y = ~ltc_cases$New_LTC.Resident.Deaths, type = 'bar', 
               name = 'Residents', 
               marker = list(color = '#AADAFF')) 

fig <- fig %>% add_trace(y = ~ltc_cases$New_LTC.HCW.Deaths, name = 'Health Care Workers', marker = list(color = '#FF5251')) 


fig <- fig %>% layout(title = "",
                      xaxis = list(title = "Time"),
                      yaxis = list (title = "New Deaths"), 
                      barmode = 'stack',
                      hovermode = "x unified")
fig
```

### <b> New Deaths per Day (Relative Change) </b>
```{r}
fig <- plot_ly(x = ~ltc_cases$Reported.Date, y = ~ltc_cases$New_LTC.Resident.Pct, type = 'bar', 
               name = 'Residents', 
               marker = list(color = '#AADAFF')) 

fig <- fig %>% add_trace(y = ~ltc_cases$New_LTC.HCW.Pct, name = 'Health Care Workers', marker = list(color = '#FF5251')) 


fig <- fig %>% layout(title = "",
                      xaxis = list(title = "Time"),
                      yaxis = list (title = "New Cases (%)"), 
                      barmode = 'stack',
                      hovermode = "x unified")
fig
```

ABOUT THE DATASET
=====================================  

### <b> ABOUT THE DATASET </b>

Dataset used was provided by the [Goverment of Ontario, Canada](https://www.ontario.ca/) through their open source data catalogue and is updated daily. 


This dataset is free to use and can be obtained [here](https://data.ontario.ca/dataset/status-of-covid-19-cases-in-ontario). Many thanks to the Government of Ontario for open sourcing this pandemic dataset. Without open sourced datasets, dashboards like this would not be  possible. 

Below is a dataset print out in case anyone is interested in the structure 

Please note some of the following fields in this dataset were intially created but after review were retired and as such no longer updated. These 'retired' fields are as follows: 

* **Confirmed Negative**	-	Patients tested negative based on testing performed at both Public Health Ontario (PHO) Laboratory and National Microbiology Lab (NML).

* **Presumptive Negative** - Patients tested negative at Public Health Ontario (PHO) Laboratory, but not yet tested by National Microbiology Lab (NML).


* **Presumptive Positive** - Patients tested positive at Public Health Ontario (PHO)  Laboratory, but not yet tested by National Microbiology Lab (NML).

```{r}
library(DT)

datatable(caseStatusOntario)
```

This dashboard was made using open sourced technologies
