---
title: "HotSpotTracker"
author: Patrick Reza Schnurbusch 
date: '`r paste("Updated: ", format(Sys.time(), "%Y-%m-%d"))`'
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    theme: paper
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)

library(geojsonio)
library(leaflet)
library(leaflet.extras)
library(leaflet.providers)
library(sf)

library(dplyr)
library(tidyr)

library(lubridate)

#read polygon data 
loc = "https://opendata.arcgis.com/datasets/d8fba69152e4408dabfe70e85a2688d2_44.geojson"
PHU_boundaries <- geojsonio::geojson_read(loc, what = "sp")

PHU_boundaries_sf <- sf::st_as_sf(PHU_boundaries)
```

GIS
===================================== 
Row {.tabset .tabset-fade}
-------------------------------------

```{r}
#download casestatus data 
CaseStatus <- read.csv(file = "https://data.ontario.ca/dataset/f4112442-bdc8-45d2-be3c-12efae72fb27/resource/455fd63b-603d-4608-8216-7d8647f43350/download/conposcovidloc.csv")
```


```{r}
CaseStatus$Case_Reported_Date <- as.Date(CaseStatus$Case_Reported_Date)

PHU_csums <- data.frame(CaseStatus %>% 
  dplyr::group_by(Reporting_PHU) %>% 
  dplyr::count(name = "csum") %>% 
  dplyr::arrange(Reporting_PHU) %>% 
  dplyr::mutate(Reporting_PHU = trimws(Reporting_PHU, which = "both"))
)

 #fixing 5 rows that the PHU name is not verbatim as in the shp boundary file 
PHU_csums$Reporting_PHU <- ifelse(PHU_csums$Reporting_PHU == "Huron Perth District Health Unit", "Huron Perth Health Unit", 
                               ifelse(PHU_csums$Reporting_PHU == "Kingston, Frontenac and Lennox & Addington Public Health", "Kingston, Frontenac and Lennox and Addington Health Unit",
                                      ifelse(PHU_csums$Reporting_PHU == "Sudbury & District Health Unit", "Sudbury and District Health Unit", 
                                             ifelse(PHU_csums$Reporting_PHU == "Wellington-Dufferin-Guelph Public Health", "Wellington-Dufferin-Guelph Health Unit", 
                                                    ifelse(PHU_csums$Reporting_PHU == "York Region Public Health Services", "York Region Public Health", PHU_csums$Reporting_PHU)
                                             )
                                      )
                               )
)
```


### <b> Cumulative per PHU </b> 
```{r}
csum_map <- merge(PHU_boundaries_sf, PHU_csums, by.x = "PHU_NAME_ENG", by.y = "Reporting_PHU" )

pal <- colorNumeric("RdBu", domain = csum_map$csum, reverse = T)

labels <- sprintf(
  "<strong>%s</strong><br/>%g Total Cases",
  csum_map$PHU_NAME_ENG, csum_map$csum
) %>% lapply(htmltools::HTML)


leaflet() %>% 
  setView(lat = 51.2538, lng = -85.3232, zoom = 5.25) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data  = csum_map,
              fillColor = ~pal(csum), 
              weight = 2, 
              opacity = 1, 
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.65, 
              
              highlightOptions = highlightOptions(
                weight = 4, 
                color = "yellow", 
                fillOpacity = 0.65, 
                bringToFront = T), 
              
              label = labels,
              
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")
  ) %>%
  
  addMiniMap(tiles = "CartoDB.Positron", position = "topright") %>% 
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  
  ## adding locate me button 
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true});}"))) %>% 
  
  addLegend("bottomright", pal = pal, values = csum_map$csum, 
            title = "Current Cumulative Cases", 
            opacity = 0.5)
```

### <b> Today's Cases </b> 

```{r}
#todays_cases 

#remove na 
CaseStatus <- CaseStatus[!(is.na(CaseStatus$Case_Reported_Date)),]

todays_cases <- data.frame(CaseStatus %>% 
                             #remove rows with an NA in 
                             dplyr::filter(Case_Reported_Date == max(Case_Reported_Date)) %>% 
                             dplyr::group_by(Reporting_PHU, Case_Reported_Date) %>% 
                             dplyr::count(name = "todays_cnt") %>% 
                             dplyr::arrange(Reporting_PHU) %>% 
                             dplyr::mutate(Reporting_PHU = trimws(Reporting_PHU, which = "both"))
                           )

#fixing 5 rows that the PHU name is not verbatim as in the shp boundary file 
todays_cases$Reporting_PHU <- ifelse(todays_cases$Reporting_PHU == "Huron Perth District Health Unit", "Huron Perth Health Unit", 
                                  ifelse(todays_cases$Reporting_PHU == "Kingston, Frontenac and Lennox & Addington Public Health", "Kingston, Frontenac and Lennox and Addington Health Unit",
                                         ifelse(todays_cases$Reporting_PHU == "Sudbury & District Health Unit", "Sudbury and District Health Unit", 
                                                ifelse(todays_cases$Reporting_PHU == "Wellington-Dufferin-Guelph Public Health", "Wellington-Dufferin-Guelph Health Unit", 
                                                       ifelse(todays_cases$Reporting_PHU == "York Region Public Health Services", "York Region Public Health", todays_cases$Reporting_PHU)
                                                )
                                         )
                                  )
)




todays_case_sf <- PHU_boundaries_sf %>% 
  dplyr::filter(PHU_boundaries_sf$PHU_NAME_ENG %in% todays_cases$Reporting_PHU)


todays_case_sf <- merge(todays_case_sf, todays_cases, by.x = "PHU_NAME_ENG", by.y = "Reporting_PHU" )
pal <- colorNumeric("RdBu", domain = todays_case_sf$todays_cnt, reverse = T)

labels <- sprintf(
  "<strong>%s</strong><br/>%g Cases(s) Today",
  todays_case_sf$PHU_NAME_ENG, todays_case_sf$todays_cnt
) %>% lapply(htmltools::HTML)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data  = todays_case_sf,
              fillColor = ~pal(todays_cnt), 
              weight = 2, 
              opacity = 1, 
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.25, 
              
              highlightOptions = highlightOptions(
                weight = 4, 
                color = "yellow", 
                fillOpacity = 0.5, 
                bringToFront = T), 
              
              label = labels,

              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")
  ) %>%
  
  addMiniMap(tiles = "CartoDB.Positron", position = "topright") %>% 
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  
  ## adding locate me button 
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true});}"))) %>% 
  
  addLegend("bottomright", pal = pal, values = todays_case_sf$todays_cnt, 
            title = "Today's Cases", 
            opacity = 0.5)
```

About this Site
===================================== 

#### Last Update  

04 September 2020 

#### Background
*"Nothing in life is to be feared, it is only to be understood. Now is the time to understand more, so that we may fear less" - Marie Curie*  


#### Code
Code and input data used to generate this Shiny mapping tool are available on [Github](https://github.com/Erebus54/OntarioCovid19TrackerDashboard/tree/master/HotSpotTracker).

#### Sources 

**Status of COVID-19 cases in Ontario** : [Open Data Ontario](https://data.ontario.ca/dataset/status-of-covid-19-cases-in-ontario), updated daily, which presents a breakdown by region, case status, age range of cases 


#### Author 
Patrick Schnurbusch 

#### <b> Contact </b> 

patrick.schnurbusch@gmail.com
